---
title: "Homework 3"
format:
  html:
    theme: [default, custom_styles]
    df-print: paged
    smaller: true
    toc: true
    toc-location: left
    toc-depth: 3
    embed-resources: true
    code-link: true
    code-tools: true
code-annotations: select
---

```{r setup, include=FALSE }
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

The code below will load replication data from: Hankinson, Michael. "When do renters behave like homeowners? High rent, price anxiety, and NIMBYism." American political science review 112.3 (2018): 473-493.

The relevant variable descriptions can be downloaded from [this site](https://dataverse.harvard.edu/file.xhtml?persistentId=doi:10.7910/DVN/CRPAA8/RSFP0V&version=1.0#) and are under the heading "TESS National Survey". (I've also included this table at the end of this document in case you can't find it.)

```{r, cache=TRUE}

library(tidyverse)
socpoc <- dataverse::get_dataframe_by_name(
  filename = "socpocAPSR.csv",
  .f=read_csv,
  dataset = "10.7910/DVN/CRPAA8&version=1.0", 
  server = "dataverse.harvard.edu")|>
  select(-1)



 
```

# Question 1

Estimate the following bivariate model using a *linear probability model*, a *logit* and a *probit* model. `ban_dummy ~ own`

## Part A

Present your results in a properly formatted table.

```{r}


```

## Part B

Based on your model, what is the effect of a home ownership on support for a ban on new home construction? State your answer in terms of the change in predicted probability of supporting a ban on new home construction for homeowners compared to renters, and present a 95% confidence interval for your estimated marginal effect.

```{r}



```

# Question 2

Estimate a linear probability and logit/probit model with the same dependent variable, but include additional controls for ideology, white (non-hispanic) and income:

`ban_dummy ~ own + ideology + whitenh + income`

```{r}


```

## Part A

Present your results from the three models in a properly formatted table and discuss any differences in model fit, significance, or the direction of effects.

## Part B

Calculate the marginal effect and a 95% confidence interval for the effect of home ownership for a white non-hispanic person whose income is \$67,500 per year at *each level* of `ideology`. Create a plot to show your estimated marginal effects and briefly discuss your results.

```{r}


```

## Part C

Estimate the average effect of homeownership across all observations using the observed values/counterfactual approach. Note any differences in your results and discuss the outcome in terms of both statistical significance and substantive importance.

```{r}

```

## Part D

The `name` variable contains the name of each respondent's home city. Estimate a fixed effects logit model (you can use the `fixest` package with the syntax `feglm(y ~ x | fe, family ='binomial')` to get a fixed effects logit model). Compare and contrast your results when using the fixed effects model. What differences do you see and what might explain these differences?

```{r}



```

## Bonus

`ban_dummy` is a dummy recoded version of `neighborhood_ban`, which was originally measured on a 7 point scale, with higher values indicating strong support. Re-estimate the model from question 2 using an ordered logit model. Then estimate the average marginal effect of homeownership on the probability of giving a response of "somewhat support", "support" or "strongly support"?

```{r}


```

# Variable Descriptions

|  |  |
|------------------|------------------------------------------------------|
| Variable | Description |
| own | “1” = Homeowner, “0” = Renter. |
| city_supply | “From your ZIP code, you live in \[INSERT CITY NAME\], which has \[INSERT TOTAL NUMBER OF UNITS BY CITY\] housing units (homes and apartments). Imagine \[INSERT CITY NAME\] lowers development restrictions, making it easier to build new housing units. As a result, \[INSERT TEN PERCENT TOTAL UNITS VALUE\] more units, with a similar mix of homes and apartments, will be built over the next five years. Would you support the lowering of development restrictions in \[INSERT CITY NAME\] to allow the construction of \[INSERT TEN PERCENT TOTAL UNITS VALUE\] more housing units over the next five years?” (1 = “Strongly Oppose”, 2 = “Oppose”, 3 = “Somewhat Oppose”, 4 = “Neutral/Uncertain”, 5 = “Somewhat Support”, 6 = “Support”, 7 = “Strongly Support”.) |
| supply_dummy | Dichotomize city_supply (“1” = “Somewhat Support”, “Support”, and “Strongly Support”; “0” = “Somewhat Oppose”, “Oppose”, “Strongly Oppose”; NA = “Neutral/Uncertain”.) |
| ideology | “In general, do you think of yourself as… ("Extremely liberal"="7", "Liberal"="6", "Slightly liberal"="5", "Moderate, middle of the road"="4", "Slightly conservative"="3", "Conservative"="2", "Extremely conservative"="1".) |
| income | Household income ("Less than \$5,000"="1", "\$5,000 to \$7,499"="2", "\$7,500 to \$9,999"="3", "\$10,000 to \$12,499"="4","\$12,500 to \$14,999"="5", "\$15,000 to \$19,999"="6", "\$20,000 to \$24,999"="7", "\$25,000 to \$29,999"="8", "\$30,000 to \$34,999"="9", "\$35,000 to \$39,999"="10", "\$40,000 to \$49,999"="11", "\$50,000 to \$59,999"="12", "\$60,000 to \$74,999"="13", "\$75,000 to \$84,999"="14", "\$85,000 to \$99,999"="15", "\$100,000 to \$124,999"="16", "\$125,000 to \$149,999"="17", "\$150,000 to \$174,999"="18", "\$175,000 or more"="19".) |
| whitenh | Dummy variable (“1” = White, Non-Hispanic, “0” = All other possibilities) |
| age | Transformed from ranges to average value of range ("18-24"="21","25-34"="29","35-44"="39","45-54"="49", "55-64"="59","65-74"="69","75+"="79")) |
| male | “1” = Male, “0” = Female. |
| name | Name of municipality. |
| zri_city | Citywide average rent, Zillow.com, June 2016. |
| zri | ZIP code average rent, Zillow.com, June 2016. |
| neighborhood_ban | “Would you support a ban on the construction of new housing (homes and apartments) in your neighborhood?” (1 = “Strongly Oppose”, 2 = “Oppose”, 3 = “Somewhat Oppose”, 4 = “Neutral/Uncertain”, 5 = “Somewhat Support”, 6 = “Support”, 7 = “Strongly Support”.) |
| ban_dummy | Dichotomize neighborhood_ban (1 = “Somewhat Support”, “Support”, and “Strongly Support”; 0 = “Somewhat Oppose”, “Oppose”, “Strongly Oppose”; NA = “Neutral/Uncertain”.) |
